---
title: "Process GC Data"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup}
library(readr)
library(magrittr)
library(purrr)
library(ggplot2)
```

This workbook is to convert peak area data from Peak Simple into PPMs. 

## Inputs

### Peak Simple files

These are the files to read in. Make sure to have saved the GC data as CSV file without headers.

```{r}
folderpath <- "../Delmarva/data/gc-peaks"
files <- list.files(folderpath, full.names = TRUE)
files
```

### Calibration curve data

This file has the PPM values for the standards. Make sure lab sheet uses matching values in `sample_type` column. 

```{r}
std_values <- read_csv("data/std_values.csv")
```

```{r, echo = FALSE, results='asis'}
std_values %>% knitr::kable()
```

### Sample IDs

Read in file with Lab data as `gc-samps`. Need columns: machineID, sample_type, and exetainer_ID.

```{r read_in_labsheet}
gc_date = "28-Oct-2018"
gc_samps <- read_csv(paste0("../Delmarva/data/gc-lab/Methane Lab Data - ",
                            gc_date, ".csv"))
```

```{r, echo = FALSE, results='asis'}
head(gc_samps) %>% knitr::kable()
```


# Read in Peak Simple data

Read in one or more Peak Simple output files saved as CSVs with `read_peaksimple()` function. This adjusts rows with 0 methane readings and adds some column names for co2 and methane columns. 

```{r}
source('read_peaksimple.R')
fileid_gc_date <- grep(gc_date, files)
gc_peaks <- read_peaksimple(files[fileid_gc_date])
```

```{r, echo = FALSE, results='asis'}
head(gc_peaks) %>% knitr::kable()
```

Add GC peak info for each row of lab data. Merges based on the numeric part of the ID from the Peak Simple file. 

```{r}
gc_data <- left_join(gc_samps, gc_peaks,
                     by = c("machineID" = "instrumentID"))
```

# Calibration curves

Use `get_stds()` function to extract the calibration standards data based on sample types matching the gases in `std_values`. This function by default saves the calibration data in `data/stds/` with the corresponding machine ID range. So can use older standard curves if needed. 

```{r}
source('get_stds.R')
ch4_stds <- get_stds(gc_data, std_values, gas = "ch4")
co2_stds <- get_stds(gc_data, std_values, gas = "co2")
# co2_stds <- read_csv("data/stds/co2_stds_2585-2655.csv")[1:7,]
```

Plot CH4 standards
```{r}
ch4_stds %>% 
  ggplot(aes(x = ppm_value, y = ch4_peak_area)) +
  geom_smooth(method = "lm") +
  geom_point(size = 3, pch = 21) + theme_bw()
```

Plot CO2 standards 
```{r}
co2_stds %>% 
  ggplot(aes(x = ppm_value, y = co2_peak_area)) +
  geom_smooth(method = "lm") +
  geom_point(size = 3, pch = 21) + theme_bw()
```

Compare curves for beginning and ending calibration standards

```{r}
ch4_calb <- lm(ch4_peak_area ~ 0 + ppm_value, data = ch4_stds)
ch4_rsquared <- summary(ch4_calb)$adj.r.squared
b1 <- ch4_calb$coefficients["(Intercept)"]
m1 <- ch4_calb$coefficients["ppm_value"]
ch4_stderror <- summary(ch4_calb)$coefficients[,2] # std errors
```


```{r}
# ch4_stds %>%
#   mutate(calbID = c(rep("cal1", 9), rep("cal2", 5))) %>%
#   ggplot(aes(x = ppm_value, y = ch4_peak_area, color = calbID)) +
#   geom_point() +
#   geom_smooth(method = "lm")
```

CO2

```{r}
# co2_stds %>%
#   mutate(calbID = c(rep("cal1", 7), rep("cal2", 6))) %>%
#   ggplot(aes(x = ppm_value, y = co2_peak_area, color = calbID)) +
#   geom_point() +
#   geom_smooth(method = "lm")
```

```{r}
co2_calb <- lm(co2_peak_area ~ 0 + ppm_value, data = co2_stds[])
co2_rsquared <- summary(co2_calb)$adj.r.squared
b2 <- co2_calb$coefficients["(Intercept)"]
m2 <- co2_calb$coefficients["ppm_value"]
co2_stderror <- summary(co2_calb)$coefficients[,2] # std errors

```


Calculate PPMs using a linear model with no intercept

```{r}
source('calculate_ppms.R')
sample_data <- calculate_ppms(gc_data, ch4_stds = ch4_stds, co2_stds = co2_stds)
# sample_data <- calculate_ppms(gc_data, ch4_stds = ch4_stds[1:9,], co2_stds = co2_stds[1:7,])
# sample_data2 <- calculate_ppms(gc_data, ch4_stds = ch4_stds[10:14,], co2_stds = co2_stds[8:13,])
```

```{r}
# sample_data[,c(1,2)] %>% rename(ch4_ppm1 = ch4_ppm) %>%
#   left_join(sample_data2[,c(1,2)]) %>% View()
#   ggplot(aes(x = ch4_ppm1, y = ch4_ppm)) +
#   geom_abline(slope = 1, intercept = 0, lty = 2) +
#   geom_point(pch = 21, size = 3) +
#   theme_bw()
```

# Save new data

```{r}
sample_file_name <- paste0("gc-ppms_", gc_date, ".csv")
write.csv(sample_data, file = file.path("results", "ppms", sample_file_name),
          row.names = FALSE)
```

# Add field info about samples

Read in field data information from all field sheets

```{r read_in_fielddata}
list.files("../Delmarva/data/field-sheets/")
# sample_info_files <- c("Methane Field Data - 11-April-2018.csv",
#                        "Methane Field Data - 12-April-2018.csv")
# sample_info_files <- c("Methane Field Data - 2018-05-02.csv")
sample_info_files <- list.files("../Delmarva/data/field-sheets/")

# read_csv(file.path("../Delmarva/data", "field-sheets", sample_info_files[27]),
#          col_types = list(Time = col_time()))

sample_info <- sample_info_files %>%
  map(function(x) read_csv(file.path("../Delmarva/data", "field-sheets", x),
                           col_types = list(Time = col_time()))) %>%
  reduce(rbind) %>% 
  dplyr::filter(!is.na(Exetainer_ID))

```

```{r}
sample_info %>%
  filter(Exetainer_ID %in% sample_info[which(duplicated(sample_info$Exetainer_ID)),"Exetainer_ID"]$Exetainer_ID)
```

Join sample info and data

```{r join_info_to_data}
# data_path <- "results/ppms/"
# files <- dir(data_path, pattern = "gc-ppms_") # get file names

# this gets all ppm data if needed
# sample_data <- files[] %>%
#   map(function(x) read_csv(file.path(data_path, x))) %>%
#   reduce(rbind)

# starting with GC data and adding site info
gc_results_db <- sample_data %>%
  left_join(sample_info, by = c("exetainer_ID" = "Exetainer_ID")) %>%
  dplyr::select(-Timepoint, -Notes, -Sample_description) %>%
  rename(Exetainer_ID = exetainer_ID)

# starting with site info data and adding GC data
# gc_results_db <- sample_info %>% 
#   left_join(sample_data, by = c("Exetainer_ID" = "exetainer_ID")) %>%
#   dplyr::select(-Timepoint, -Notes, -Sample_description)

# gc_results_db %>%
#   mutate(mr = Exetainer_ID %in% sample_data$exetainer_ID)
  
# sample_data %>% left_join(sample_info, by = c("exetainer_ID" = "Exetainer_ID")) %>% View()

# length(which(is.na(gc_results_db_sub$Field_sample_date)))

gc_results_db_sub <- gc_results_db %>% filter(Exetainer_ID %in% sample_data$exetainer_ID)

# write.csv(gc_results_db, "results/ppms_wsampleinfo/gc_results_08-09-March-2018.csv", row.names = FALSE)
gc_results_filename <- paste0("gc_results_", gc_date, ".csv")
write.csv(gc_results_db_sub, 
          file = file.path("results", "ppms_wsampleinfo", gc_results_filename),
          row.names = FALSE)

```

# Insert PPMs into database

Use current-gas-samples.Rmd to figure out which exetainer IDs have results in the database already, based on the samplecode column in `gas_spread`. If this returns nothing then none of the samples in gc_results_db_sub have results in the database already. 

```{r}
gc_results_db_sub$Exetainer_ID[gc_results_db_sub$Exetainer_ID %in% gas_spread$samplecode]
```

Make datetime column

```{r}
gc_results_db_sub$Timestamp <- as.POSIXct(strptime(paste(gc_results_db_sub$Field_sample_date,
                                             gc_results_db_sub$Time),
                                       format = "%B %d, %Y %H:%M:%S"))
```

```{r}
gc_results_db_sub$Timestamp_analysis <-
  as.POSIXct(strptime(paste(gc_results_db_sub$gc_date,
                                             gc_results_db_sub$gc_time),
                                       format = "%m/%d/%Y %H:%M:%S"))
```

Will need to insert one method at a time

## Soil chamber data 

Make column for site codes that match sampling feature names in database

```{r}
sc_db <- gc_results_db_sub %>%
  filter(Method == "SC") %>%
  mutate(site_code = paste(Site, Chamber_ID))
sc_db$Site <- gsub("SC ", "SC-", sc_db$site_code)
```

Check that all site names are in Database

```{r}
library(rodm2)
check_samplingfeaturecodes(sc_db$Site, db)
```

```{r}
vars <- dbGetQuery(db, "SELECT * FROM odm2.variables")
dbmethods <- dbGetQuery(db, "SELECT * FROM odm2.methods")
```

```{r}
names(sc_db)
vars
dbunits <- get_cv_terms("units")
grep("Million", dbunits, value = TRUE)
get_cv_terms("medium")
vars_list <- list("Methane" = list(column = "ch4_ppm", units = "Part per Million"),
                  "Carbon dioxide" = list(column = "co2_ppm", units = "Part per Million"))
sc_db <- sc_db %>% dplyr::select(Timestamp, Timestamp_analysis, ch4_ppm, co2_ppm, Site, Exetainer_ID)
sc_db$Timestamp <- as.character(sc_db$Timestamp)
sc_db$Timestamp_analysis <- as.character(sc_db$Timestamp_analysis)
```

```{r}
rodm2::db_insert_results_samples(db = db,
                          datavalues = sc_db,
                          field_method = "SC",
                          lab_method = "GC",
                          variables = vars_list, field_actionby = "Kelly", lab_actionby = "Kelly",
                          sampledmedium = "Gas", site_code_col = "Site", sample_code_col = "Exetainer_ID")
```

## Headspace data

```{r}
hs_db <- gc_results_db_sub %>%
  filter(Method == "HS") %>%
  dplyr::select(Timestamp, Timestamp_analysis, Exetainer_ID, Site, ch4_ppm, co2_ppm)
```

Tibble datetime format is annoying and doesn't work

```{r}
hs_db$Timestamp <- as.character(hs_db$Timestamp)
hs_db$Timestamp_analysis <- as.character(hs_db$Timestamp_analysis)
```

Check site codes against sites in database (this should return no rows)

```{r}
check_samplingfeaturecodes(hs_db$Site, db)
```

```{r}
rodm2::db_insert_results_samples(db = db,
                          datavalues = hs_db,
                          field_method = "HS",
                          lab_method = "GC",
                          variables = vars_list, 
                          field_actionby = "Kelly", lab_actionby = "Kelly",
                          sampledmedium = "Liquid aqueous", 
                          site_code_col = "Site", sample_code_col = "Exetainer_ID")
```

## Air samples

```{r}
air_db <- gc_results_db_sub %>%
  filter(Method == "air") %>%
  dplyr::select(Timestamp, Timestamp_analysis, Exetainer_ID, Site, ch4_ppm, co2_ppm)
```

```{r}
air_db$Timestamp <- as.character(air_db$Timestamp)
air_db$Timestamp_analysis <- as.character(air_db$Timestamp_analysis)
check_samplingfeaturecodes(air_db$Site, db)
```

```{r}
rodm2::db_insert_results_samples(db = db,
                          datavalues = air_db,
                          field_method = "air",
                          lab_method = "GC",
                          variables = vars_list, 
                          field_actionby = "Kelly", lab_actionby = "Kelly",
                          sampledmedium = "Air", 
                          site_code_col = "Site", sample_code_col = "Exetainer_ID")
```
