---
title: "Process GC Data"
output: 
  html_notebook:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(magrittr)
```

This workbook is to convert peak area data from Peak Simple into PPMs. 

## Inputs

### Peak Simple files

These are the files to read in. Make sure to have saved the GC data as CSV file without headers.

Set up link between folder where the GC output data lives on your computer and this project directory with a `sym.link()`.

```{r}
file.symlink("../Delmarva/data/gc-peaks", "gc-peaks") 
files <- list.files("gc-peaks", full.names = TRUE)
```

### Calibration curve data

This file has the PPM values for the standards. Make sure lab sheet uses matching values in `sample_type` column. 

```{r}
std_values <- read_csv("data/std_values.csv")
```

```{r, echo = FALSE, results='asis'}
std_values %>% knitr::kable()
```

### Sample IDs

Set up a symbolic link to the folder where lab data sheets are
```{r}
file.symlink("../Delmarva/data/gc-lab/", "gc-lab/")
```

Read in file with Lab data as `gc-samps`. Need columns: machineID, sample_type, and exetainer_ID.

```{r}
gc_date = "28-July-2018"
gc_samps <- read_csv(paste0("gc-lab/Methane Lab Data - ",
                            gc_date, ".csv"))
```

```{r, echo = FALSE, results='asis'}
head(gc_samps) %>% knitr::kable()
```


# Read in Peak Simple data

Read in one or more Peak Simple output files saved as CSVs with `read_peaksimple()` function. This adjusts rows with 0 methane readings and adds some column names for co2 and methane columns. 

```{r, message=FALSE}
source('read_peaksimple.R')
gc_peaks <- read_peaksimple(files)
```

```{r, echo = FALSE, results='asis'}
head(gc_peaks) %>% knitr::kable()
```

Add GC peak info for each row of lab data. Merges based on the numeric part of the ID from the Peak Simple file. 

```{r}
gc_data <- left_join(gc_samps, gc_peaks,
                     by = c("machineID" = "instrumentID"))
```

# Calibration curves

Use `get_stds()` function to extract the calibration standards data based on sample types matching the gases in `std_values`. This function by default saves the calibration data in `data/stds/` with the corresponding machine ID range. So can use older standard curves if needed. 

```{r}
source('get_stds.R')
ch4_stds <- get_stds(gc_data, std_values, gas = "ch4")
co2_stds <- get_stds(gc_data, std_values, gas = "co2")
```

Calculate PPMs using a linear model with no intercept

```{r}
source('calculate_ppms.R')
sample_data <- calculate_ppms(gc_data, ch4_stds, co2_stds)
```

# Save new data

```{r}
sample_file_name <- paste0("gc-ppms_", gc_date, ".csv")

if(!dir.exists(file.path("results", "ppms"))){
  dir.create("results")
  dir.create(file.path("results", "ppms"))
    }

write.csv(sample_data, file = file.path("results", "ppms", sample_file_name),
          row.names = FALSE)
```

