---
title: "Process GC Data"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(readr)
library(magrittr)
library(purrr)
library(ggplot2)
```


# Insert PPMs into database

This starts with the data that is saved in results/ppms_wsampleinfo. 

Use `current-gas-samples.Rmd` (in methane-source-area project) to figure out which exetainer IDs have results in the database already, based on the samplecode column in `gas_spread`. If this returns nothing then none of the samples in gc_results_db_sub have results in the database already. 

```{r}
# read in gc_results_db_sub if not in environment
# ie if you didn't upload right away like you were supposed to
list.files("../gc/results/ppms_wsampleinfo/", full.names = TRUE)
files <- list.files("../gc/results/ppms_wsampleinfo/", full.names = TRUE)
gc_results_db_sub <- read_csv(files[2])
```

```{r db_check_if_exists}
gc_results_db_sub$Exetainer_ID[gc_results_db_sub$Exetainer_ID %in% gas_spread$samplecode]
```

If needed, take one out for now
```{r}
gc_results_db_sub <- dplyr::filter(gc_results_db_sub, Exetainer_ID != "KLH 1055")
gc_results_db_sub <- dplyr::filter(gc_results_db_sub, Exetainer_ID != "KLH 1726")
```

Make datetime column

```{r}
gc_results_db_sub$Timestamp <- as.POSIXct(strptime(paste(gc_results_db_sub$Field_sample_date,
                                             gc_results_db_sub$Time),
                                       format = "%B %d, %Y %H:%M:%S"))
```

```{r}
gc_results_db_sub$Timestamp_analysis <-
  as.POSIXct(strptime(paste(gc_results_db_sub$gc_date,
                                             gc_results_db_sub$gc_time),
                                       format = "%m/%d/%Y %H:%M:%S"))
```

Will need to insert one method at a time

First fix site names and dates if needed.

```{r}
unique(gc_results_db_sub$Site)
unique(gc_results_db_sub$Field_sample_date)
```


## Soil chamber data 

Make column for site codes that match sampling feature names in database

```{r}
sc_db <- gc_results_db_sub %>%
  filter(Method == "SC") %>%
  mutate(site_code = paste(Site, Chamber_ID))
sc_db$Site <- gsub("SC ", "SC-", sc_db$site_code)
```

Check that all site names are in Database

```{r}
library(rodm2)
check_samplingfeaturecodes(sc_db$Site, db)
```

```{r}
vars <- dbGetQuery(db, "SELECT * FROM odm2.variables")
dbmethods <- dbGetQuery(db, "SELECT * FROM odm2.methods")
```

```{r}
names(sc_db)
vars
# dbunits <- get_cv_terms("units")
# grep("Million", dbunits, value = TRUE)
# get_cv_terms("medium")
vars_list <- list("Methane" = list(column = "ch4_ppm", units = "Part per Million"),
                  "Carbon dioxide" = list(column = "co2_ppm", units = "Part per Million"))
sc_db <- sc_db %>% dplyr::select(Timestamp, Timestamp_analysis, ch4_ppm, co2_ppm, Site, Exetainer_ID)
sc_db$Timestamp <- as.character(sc_db$Timestamp)
sc_db$Timestamp_analysis <- as.character(sc_db$Timestamp_analysis)
```

```{r}
rodm2::db_insert_results_samples(db = db,
                          datavalues = sc_db,
                          field_method = "SC",
                          lab_method = "GC",
                          variables = vars_list, field_actionby = "Kelly", lab_actionby = "Kelly",
                          sampledmedium = "Gas", site_code_col = "Site", sample_code_col = "Exetainer_ID")
```

## Headspace data

```{r}
hs_db <- gc_results_db_sub %>%
  filter(Method == "HS") %>%
  dplyr::select(Timestamp, Timestamp_analysis, Exetainer_ID, Site, ch4_ppm, co2_ppm)
```

Tibble datetime format is annoying and doesn't work

```{r}
hs_db$Timestamp <- as.character(hs_db$Timestamp)
hs_db$Timestamp_analysis <- as.character(hs_db$Timestamp_analysis)
```

Check site codes against sites in database (this should return no rows)

```{r}
check_samplingfeaturecodes(hs_db$Site, db)
```

```{r}
rodm2::db_insert_results_samples(db = db,
                          datavalues = hs_db,
                          field_method = "HS",
                          lab_method = "GC",
                          variables = vars_list, 
                          field_actionby = "Kelly", lab_actionby = "Kelly",
                          sampledmedium = "Liquid aqueous", 
                          site_code_col = "Site", sample_code_col = "Exetainer_ID")
```

## Air samples

```{r}
air_db <- gc_results_db_sub %>%
  filter(Method == "air") %>%
  dplyr::select(Timestamp, Timestamp_analysis, Exetainer_ID, Site, ch4_ppm, co2_ppm)
```

```{r}
air_db$Timestamp <- as.character(air_db$Timestamp)
air_db$Timestamp_analysis <- as.character(air_db$Timestamp_analysis)
check_samplingfeaturecodes(air_db$Site, db)
```

```{r}
rodm2::db_insert_results_samples(db = db,
                          datavalues = air_db,
                          field_method = "air",
                          lab_method = "GC",
                          variables = vars_list, 
                          field_actionby = "Kelly", lab_actionby = "Kelly",
                          sampledmedium = "Air", 
                          site_code_col = "Site", sample_code_col = "Exetainer_ID")
```

```{r}
dbDisconnect(db)
```

